---
title: "Examining the relationship between the big-5 personality facets and implicit racial attitudes"
subtitle: "Analysis"
author: "Sarah Schnyder"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---
```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(ggdist)
library(ggrain)
library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(scales)
library(ggExtra)
library(patchwork)
library(ggrepel)
library(remotes)
library(ggmagnify)
library(report)
library(psych)

```

# Data

Load the processed data and apply the global exclusions.

```{r}

data_processed <- read_csv("../data/processed/data_processed.csv")

data_processed_after_exclusions <- data_processed |>
  filter(exclusion == "include")

```

# Sample descriptives

## Sample size before exclusions

```{r}

data_processed |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("Whole sample" = 1)) |> # note that you can add header rows to tables like this. The "1" indicates the number of columns the header should span. The sum of these numbers must equal the number of columns or you'll get an error.
  kable_classic(full_width = FALSE)

```

## Sample size after exclusions

```{r}

data_processed_after_exclusions |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("For analysis" = 1)) |>
  kable_classic(full_width = FALSE)

```

## Age

```{r}

data_processed_after_exclusions |>
  mutate(age = as.numeric(age)) |>
  summarise(Mean = mean(age, na.rm = TRUE),
            SD = sd(age, na.rm = TRUE)) |>
  mutate_all(.funs = janitor::round_half_up, digits = 1) |>
  kable() |>
  add_header_above(header = c("Age" = 2)) |>
  kable_classic(full_width = FALSE)

```

## Sex

```{r}

data_processed_after_exclusions |> 
  group_by(sex) |> 
  summarise(n = n()) |> 
  mutate(
    sex = recode(sex, "f" = "female", "m" = "male"),
    Percent = paste0(round_half_up((n / sum(n)) * 100, 1), "%")
  ) |> 
  rename(Sex = sex) |> 
  kable() |> 
  kable_classic(full_width = FALSE)


```

### Availability of demographics variables
```{r}
proportion_available_age <- 
  (sum(data_processed_after_exclusions$age != "other/missing/error") / nrow(data_processed_after_exclusions)) |> 
  round(3)

proportion_available_sex <- 
  (sum(data_processed_after_exclusions$sex %in% c("f", "m")) / nrow(data_processed_after_exclusions)) |> 
  round(3)

print(paste0("Variable age was available for ", proportion_available_age, " of participants and variable sex for ",
             proportion_available_sex, " of participants."))


```

## Cronbachs alpha BFI subscales

```{r}
# defining a function to get all columns of one subscale in a vector
function_items_subscale_bfi <- function(subscale, n){
  items_subscale_bfi <- c()
  for (i in 1:n){
    items_subscale_bfi <- append(items_subscale_bfi, paste0("bfi_", subscale, i))
  }
  items_subscale_bfi
}

# agreeableness
alpha(subset(data_processed_after_exclusions, 
             select = function_items_subscale_bfi("a", 9)), check.keys =TRUE)

# conscientiousness
alpha(subset(data_processed_after_exclusions, 
             select = function_items_subscale_bfi("c", 9)), check.keys =TRUE)

# extraversion
alpha(subset(data_processed_after_exclusions, 
             select = function_items_subscale_bfi("e", 8)), check.keys =TRUE)

# neuroticism
alpha(subset(data_processed_after_exclusions, 
             select = function_items_subscale_bfi("n", 8)), check.keys =TRUE)

# openness
alpha(subset(data_processed_after_exclusions, 
             select = function_items_subscale_bfi("o", 10)), check.keys =TRUE)

```
These are the Cronbach's alpha for all the BFI subscales:

* agreeableness: 0.78
* conscientiousness: 0.81
* extraversion: 0.82
* neuroticism: 0.78
* openness: 0.81

# Descriptive Plots

## BFI subscales
```{r}

# defining a function for the histograms for the BFI subscales
histogram_bfi_subscales <- function(column, x_label){
  
  ggplot(data_processed_after_exclusions, aes(x = get(column))) +
    geom_histogram(binwidth = 0.1,
                   boundary = 0,
                   fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                   color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
    xlab(x_label) +
    ylab("Frequency") +
    coord_cartesian(xlim = c(1,6))+
    theme_linedraw() +
    scale_x_continuous(breaks = pretty_breaks(n = 10))+
    theme_minimal()
}

```

### Agreeableness
```{r}
histogram_bfi_subscales("mean_bfi_a", "Mean Agreeableness")

```



### Conscientiousness
```{r}
histogram_bfi_subscales("mean_bfi_c", "Mean Conscientiousness")


```


### Extraversion
```{r}
histogram_bfi_subscales("mean_bfi_e", "Mean Extraversion")

```

### Neuroticism
```{r}

histogram_bfi_subscales("mean_bfi_n", "Mean Neuroticism")

```

### Openness
```{r}

histogram_bfi_subscales("mean_bfi_o", "Mean Openness")

```



## IAT score
```{r}

ggplot(data_processed_after_exclusions, aes(x = IAT_score)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("IAT Score") +
  ylab("Frequency") +
  coord_cartesian(xlim = c(-2,2))+
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 10))+
  theme_minimal()

```

# Correlation Matrix

```{r}
# making a new data frame with only the columnes required for the correlatio matrix
data_processed_cols_correlation <- data_processed |> 
  select(c(mean_bfi_a, mean_bfi_c, mean_bfi_e, mean_bfi_n, mean_bfi_o, IAT_score))

# making the correlation matrix
cor(data_processed_cols_correlation, use = "pairwise.complete.obs", method="pearson") |>
  round(2)
```
Some correlations can't be computed because there are no subjects that answered that combination of subscales. E.g. there was no subject that answered the items to agreeableness and to extraversion. 


# Hypothesis Testing
## Hypothesis 1: Men and women differ on their scores on the IAT.
To test this hypothesis I'm doing a t-test.
```{r}
# to run this test I'm creating a data frame with only the subjects that gave information about their sex
data_processed_after_exclusions_sex <- data_processed_after_exclusions |> 
  filter(sex %in% c("f", "m"))

# running the t-test
t.test(IAT_score~sex, data = data_processed_after_exclusions_sex, alternative = "two.sided")

# reporting the t-test
report(t.test(IAT_score~sex, data = data_processed_after_exclusions_sex, alternative = "two.sided"))


```


## Hypothesis 2: Each BFI subscale predicts IAT scores.
To test this I'm doing a regression for each BFI subscale.



### Hypothesis 2.1: The BFI subscale agreeableness predicts IAT scores.

```{r}

# creating the model
model_a <- lm(IAT_score ~ mean_bfi_a, data = data_processed)

# get summary of the model
summary(model_a)

# reporting the model
report(model_a)

```

### Hypothesis 2.2: The BFI subscale conscientiousness predicts IAT scores.

```{r}

# creating the model
model_c <- lm(IAT_score ~ mean_bfi_c, data = data_processed)

# get summary of the model
summary(model_c)

# reporting the model
report(model_c)

```

### Hypothesis 2.3: The BFI subscale extraversion predicts IAT scores.

```{r}

# creating the model
model_e <- lm(IAT_score ~ mean_bfi_e, data = data_processed)

# get summary of the model
summary(model_e)

# reporting the model
report(model_e)


```


### Hypothesis 2.4: The BFI subscale neuroticism predicts IAT scores.

```{r}

# creating the model
model_n <- lm(IAT_score ~ mean_bfi_n, data = data_processed)

# get summary of the model
summary(model_n)

# reporting the model
report(model_n)


```

### Hypothesis 2.5: The BFI subscale openness predicts IAT scores.

```{r}

# creating the model
model_o <- lm(IAT_score ~ mean_bfi_o, data = data_processed)

# get summary of the model
summary(model_o)

# reporting the model
report(model_o)

```
# Scatter plots of how BFI subscale scores predict IAT

```{r}
# first lets make a new column defining if the IAT score is an extreme score or not
IAT_top_10_percentile_threshold <- quantile(data_processed_after_exclusions$IAT_score, 0.9, na.rm = TRUE)
IAT_bottom_10_percentile_threshold <- quantile(data_processed_after_exclusions$IAT_score, 0.1, na.rm = TRUE)

data_processed_after_exclusions_extreme_IAT <- 
  data_processed_after_exclusions |> 
  mutate(extreme_IAT = ifelse(IAT_score >= IAT_top_10_percentile_threshold | IAT_score <= IAT_bottom_10_percentile_threshold,
                              "extreme",
                              "normal"))

# lets define a function for making the plots
function_plot_BFI_IAT <- function(column, name_subscale){
  ggplot(data_processed_after_exclusions_extreme_IAT, # define which data frame should be used
         aes(x = get(column), y = IAT_score)) + # define which variable is on the x and which is on the y axis
    geom_point(alpha = 0.7, # turning down alpha a bit so you can see where are more points and where are less
               aes(shape = extreme_IAT)) + # using different shapes based on extremeness of IAT score
    scale_shape_manual(values = c(17, 16)) + # defining which shapes should be used
    geom_smooth(method = "lm") + # add regression line
    ggtitle(name_subscale)+ # add title for the whole plot
    xlab(paste0("Mean ", name_subscale))+ # add label for x axis
    ylab("IAT Score")+ # add label for y axis
    theme_classic() + # give plot a classic theme
    theme(plot.title = element_text(hjust = 0.5),  # Center the title horizontally
          legend.position = "none")  # remove the legend
  
}

```


```{r}
# defining all the plots
p_a <- function_plot_BFI_IAT("mean_bfi_a", "Agreeableness")
p_c <- function_plot_BFI_IAT("mean_bfi_c", "Conscientiousness")
p_e <- function_plot_BFI_IAT("mean_bfi_e", "Extraversion")
p_n <- function_plot_BFI_IAT("mean_bfi_n", "Neuroticism")
p_o <- function_plot_BFI_IAT("mean_bfi_o", "Openness")
 

# combine all plots into one
combined_plot <- p_a + p_c + p_e + p_n + p_o 
print(combined_plot)

# save the combined plot
ggsave("../communications/combined_plot.png", width = 20, plot = combined_plot, dpi = 300, device = "png")

ggsave("../communications/combined_plot.pdf", width = 20, plot = combined_plot, dpi = 300, device = "pdf")
```




```{r}
sessionInfo()
```

