---
title: "Examining the relationship between the big-5 personality facets and implicit racial attitudes"
subtitle: "Processing"
author: "Sarah Schnyder"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# Dependencies

```{r}

library(tidyverse)
library(janitor) 
library(stringr)
library(openxlsx)
library(kableExtra)

```


# Get data

```{r}

data_bfi_raw <- read_csv("../data/raw/data_raw_bfi.csv") |>
  janitor::clean_names()

data_iat_raw <- read_csv("../data/raw/data_raw_iat.csv", skip = 1) |> # skip = 1, because the column names are in the second row
  janitor::clean_names()

data_demographics_raw <- read_csv("../data/raw/data_raw_demographics.csv") |>
  janitor::clean_names()

```

# Demographics
## Extract age and gender from raw demographics data

```{r}

# remove the rows that have missing data in the column unique_id.
# These rows are unusable, because I can't match them with other data.
data_demographics_without_missing_ids <-
  data_demographics_raw |> 
  filter(!is.na(unique_id))

# transform data into wide format and check if the data is correct
data_demographics_wide <- data_demographics_without_missing_ids |> 
  pivot_wider(
    names_from = variable,
    values_from = response
  ) |> 
  mutate(sex = case_when(sex %in% c("f", "m") ~ sex,
                         TRUE ~ "missing/other/error"),
         age = case_when(str_detect(age, "^[0-9]+$") ~ age, # if the value is only numbers, keep it. 
                         TRUE ~ "other/missing/error")
  )

```


# BFI
## Reverse score the negatively worded items
The negatively worded items are: the extroversion scale items 2, 5 and 7, conscientiousness items 2, 4 5 and 9, neuroticism items 2, 5, and 7, agreeableness 1, 3, 6, and 8, and openness items 7 and 9.

```{r}

# make vector with all columns, that need to be reverse scored
cols_reverse_score <- c("e2", "e5", "e7",
                        "c2", "c4", "c5", "c9",
                        "n2", "n5", "n7",
                        "a1", "a3", "a6", "a8",
                        "o7", "o9")
cols_reverse_score <- paste0("bfi_", cols_reverse_score)

max_possible_score <- 6 + 1


data_bfi_reverse_scored <- data_bfi_raw |> 
  mutate(across(cols_reverse_score, ~ max_possible_score - .))


```

### Sanity check for reverse scoring
For each subscale, create a correlation table among the items (after reversals) and check that all correlations are positive. IN general, negative correlations among items are one indication that information about item reversals, or their implementation, is not correct.
```{r}


for(i in c("a", "c", "e", "n", "o")){
  data_bfi_reverse_scored_subscale <- data_bfi_reverse_scored %>%
    select(contains(paste0("bfi_", i))) # selecting only the columns of one subscale
  
  correlation_matrix <- cor(data_bfi_reverse_scored_subscale, use = "complete.obs") # making a correlation matrix
  print(correlation_matrix)
  print(ifelse(all(correlation_matrix > 0), 
               paste0("All the correlations of the subscale ",
                      i, " are positive."), 
               paste0("NOT all of the correlations of the subscale ",
                      i, " are positive."))) # checking if all the correlations are positive
}



```
The only negative correlation found is between bfi_o7 and bfi_010. But since it's only one and that one is relatively small, I can take this check as a confirmation of the correctness of my reverse scoring.

## BFI Excluding participants 

Criteria for excluding participants:

* values are not between 1 and 6
* incomplete data on completed BFI scales

```{r}

# set new variable to include if all the values are between 1 and 6 or are missing value
# set another new variable to include if either all values of one subscale or missing or all values of one subscale are not missing
data_bfi_reverse_scored_exclusion_variable <- data_bfi_reverse_scored |> 
  rowwise() |> 
  mutate(exclude_between_1_and_6 = ifelse(
    all(c_across(starts_with("bfi_")) %in% 1:6 | is.na(c_across(starts_with("bfi_")))), 
    "include",
    "exclude"
  ),
  exclude_incomplete_data = ifelse(
    (all(is.na(c_across(starts_with("bfi_a")))) | all(!is.na(c_across(starts_with("bfi_a"))))) &
      (all(is.na(c_across(starts_with("bfi_c")))) | all(!is.na(c_across(starts_with("bfi_c"))))) &
      (all(is.na(c_across(starts_with("bfi_e")))) | all(!is.na(c_across(starts_with("bfi_e"))))) &
      (all(is.na(c_across(starts_with("bfi_n")))) | all(!is.na(c_across(starts_with("bfi_n"))))) &
      (all(is.na(c_across(starts_with("bfi_o")))) | all(!is.na(c_across(starts_with("bfi_o"))))),
    "include",
    "exclude"
  )
  )


```


## Mean-score of subscales

```{r}


data_bfi_reverse_scored_exclusion_variable$mean_bfi_a <- rowMeans(select(data_bfi_reverse_scored_exclusion_variable,starts_with("bfi_a")),
                                                                  na.rm = TRUE)
data_bfi_reverse_scored_exclusion_variable$mean_bfi_c <- rowMeans(select(data_bfi_reverse_scored_exclusion_variable,starts_with("bfi_c")),
                                                                  na.rm = TRUE)
data_bfi_reverse_scored_exclusion_variable$mean_bfi_e <- rowMeans(select(data_bfi_reverse_scored_exclusion_variable,starts_with("bfi_e")),
                                                                  na.rm = TRUE)
data_bfi_reverse_scored_exclusion_variable$mean_bfi_n <- rowMeans(select(data_bfi_reverse_scored_exclusion_variable,starts_with("bfi_n")),
                                                                  na.rm = TRUE)
data_bfi_reverse_scored_exclusion_variable$mean_bfi_o <- rowMeans(select(data_bfi_reverse_scored_exclusion_variable,starts_with("bfi_o")),
                                                                  na.rm = TRUE)

# check if the means are not smaller than 1 and not bigger than 6

for (i in c("a", "c", "e", "n", "o")){
  if(data_bfi_reverse_scored_exclusion_variable |> 
     mutate(bounded_correctly = between(get(paste0("mean_bfi_", i)), left = 1, right = 6)) |>
     filter(bounded_correctly != TRUE) |>
     nrow() == 0) {
    print(paste0("The mean score of the subscale ", i, " does not violate the min and max possible score."))
  } else {
    print(paste0("The mean score of the subscale ", i, " does violate the min and max possible score."))
  }
}


```
None of the mean score violates the min and max possible score. So I can take this as one hint, that the mean was build correctly.


# IAT
## Scoring the trial-level IAT

Greenwald "D" score: 

* Calculate a mean RT ("mean1") for blocks 3 and 6 (one score using trials from both blocks)
* a mean RT ("mean2") for blocks 4 and 7 (one score using trials from both blocks)
* and the SD of RTs in blocks 3, 4, 6 and 7 ("SD")
* To calculate D: D = (mean2 - mean1)/SD
* Blocks 1, 2, and 5 are practice blocks and must be discarded.

```{r}

# removing practice blocks
data_iat_without_practice <-
  data_iat_raw |> 
  filter(!block_number %in% c(1,2,5))

# Group by unique_id
data_iat_without_practice_grouped <- data_iat_without_practice %>%
  group_by(unique_id)

# Calculating mean RT for blocks 3 and 6 for each unique_id
data_iat_without_practice_grouped_means <- data_iat_without_practice_grouped %>%
  filter(block_number %in% c(3, 6)) %>%
  summarise(mean1 = mean(trial_reaction_time_in_ms))

# Calculate mean RT for blocks 4 and 7 for each unique_id
data_iat_without_practice_grouped_means<- data_iat_without_practice_grouped %>%
  filter(block_number %in% c(4, 7)) %>%
  summarise(mean2 = mean(trial_reaction_time_in_ms)) %>%
  left_join(data_iat_without_practice_grouped_means, by = "unique_id")

# Calculate SD of RTs for blocks 3, 4, 6, and 7 for each unique_id
data_iat_without_practice_grouped_sd <- data_iat_without_practice_grouped %>%
  filter(block_number %in% c(3, 4, 6, 7)) %>%
  summarise(SD = sd(trial_reaction_time_in_ms))

# Calculate D for each unique_id
data_iat_without_practice_D_score <- data_iat_without_practice_grouped_means %>%
  left_join(data_iat_without_practice_grouped_sd, by = "unique_id") %>%
  mutate(IAT_score = (mean2 - mean1) / SD) |> 
  select(unique_id, IAT_score) |> 
  ungroup()

# Print the new df
print(data_iat_without_practice_D_score)


# sanity check: Are all D scores in the range -2 to +2?
if (all(data_iat_without_practice_D_score$IAT_score >= -2 & data_iat_without_practice_D_score$IAT_score <= 2)) {
  print("All D scores are in the range -2 to +2.")
} else {
  print("Some D scores are outside the range -2 to +2.")
}


```

## Exclusions
Criteria for exclusions are:

* incomplete trial level IAT data (normal are 120 trials)
* \>10% of the trials are < 300 ms
* accuracy is < 75%

```{r}
# getting a vector with all the IDs that don't have 120 trials
iat_exclusion_incomplete_trial <- data_iat_without_practice |> 
  group_by(unique_id) |> 
  filter(all(c(3, 4, 6, 7) %in% block_number) & n() != 120) |> 
  pull(unique_id) |> 
  unique()

# getting a vector with all the Ids that have < 300 ms reactions times in more than 10% of the trials
data_iat_without_practice_rt <-
  data_iat_without_practice |> 
  mutate(short_rt = ifelse(trial_reaction_time_in_ms < 300, "short", "ok"))

iat_exclusion_short_rt <- data_iat_without_practice_rt |> 
  group_by(unique_id) |> 
  mutate(proportion_short_rt = sum(short_rt == "short")/ n()) |> 
  filter(proportion_short_rt > 0.10) |> 
  pull(unique_id) |> 
  unique()

# getting a vector with all the Ids that have an accuracy < 75%
iat_exclusion_low_accuracy <- data_iat_without_practice |> 
  group_by(unique_id) |> 
  mutate(accuracy = sum(trial_accuracy == "correct")/ n()) |> 
  filter(accuracy < 0.75) |> 
  pull(unique_id) |> 
  unique()

# create one vector with all the Ids that meet at one of the exclusion criteria
iat_exclusion <- c(iat_exclusion_incomplete_trial, iat_exclusion_short_rt, iat_exclusion_low_accuracy)


data_iat_without_practice_D_exclusions <-
  data_iat_without_practice_D_score |> 
  mutate(exclude_iat = case_when(unique_id %in% iat_exclusion ~ "exclude",
                                 TRUE ~ "include"))

```

# Combining demographics, BFI and IAT 

Creating a one-row-one-participant data frame out of all of the data frames.

```{r}
data_processed <-
  data_demographics_wide |>
  full_join(data_bfi_reverse_scored_exclusion_variable, by = "unique_id") |> 
  full_join(data_iat_without_practice_D_exclusions, by = "unique_id") 

print(data_processed)
```

## Master exclude variable

```{r}
data_processed_master_exclude <-
  data_processed |> 
  mutate(
    exclusion = case_when(exclude_between_1_and_6 == "exclude" ~ "exclude",
                          exclude_incomplete_data == "exclude"  ~"exclude",
                          exclude_iat == "exclude" ~"exclude",
                          TRUE ~ "include")
  ) |> 
  select(-c(exclude_between_1_and_6, exclude_incomplete_data, exclude_iat))


```



# Write to disk
```{r}
# in case this dir doesn't exist, create it
dir.create("../data/processed/")

# save data to disk in that dir
write_csv(data_processed_master_exclude, "../data/processed/data_processed.csv")

```


# Create Codebook
```{r}
if(!file.exists("../data/processed/data_processed_codebook.xlsx")){
  # convert the column names to a df
  codebook_template <- data.frame(variable = colnames(data_processed_master_exclude)) |>
    mutate(explanation = NA)
  # write to disk as an excel file
  write.xlsx(codebook_template, file = "../data/processed/data_processed_codebook.xlsx")
}
```

