---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Analysis"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(ggdist)
library(ggrain)
library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(scales)
library(ggExtra)
library(patchwork)
library(ggrepel)
library(remotes)
library(ggmagnify)
library(report)


```

# Data

Load the processed data and apply the global exclusions.

```{r}

data_processed <- read_csv("../data/processed/data_processed.csv")

data_processed_after_exclusions <- data_processed |>
  filter(exclude_participant == "include")

```

# Sample descriptives

## Sample size before exclusions

```{r}

data_processed |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("Whole sample" = 1)) |> # note that you can add header rows to tables like this. The "1" indicates the number of columns the header should span. The sum of these numbers must equal the number of columns or you'll get an error.
  kable_classic(full_width = FALSE)

```

## Sample size after exclusions

Sample used in subsequent analyses

```{r}

data_processed_after_exclusions |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("For analysis" = 1)) |>
  kable_classic(full_width = FALSE)

```

## Age

```{r}

data_processed_after_exclusions |>
  mutate(age = as.numeric(age)) |>
  summarise(Mean = mean(age, na.rm = TRUE),
            SD = sd(age, na.rm = TRUE)) |>
  mutate_all(.funs = janitor::round_half_up, digits = 1) |>
  kable() |>
  add_header_above(header = c("Age" = 2)) |>
  kable_classic(full_width = FALSE)

```

## Gender

```{r}

data_processed_after_exclusions |> 
  rename(Gender = gender) |>
  group_by(Gender) |> 
  summarise(n = n()) |> 
  mutate(Percent = paste0(round_half_up((n / sum(n)) * 100, 1), "%")) |>
  mutate(Gender = stringr::str_to_sentence(Gender)) |> # Change the case of the Gender variable so that it prints nicely
  kable() |>
  kable_classic(full_width = FALSE)

```

# Descriptives

Descriptive statistics and plots of the measures (excluding the demographics variables)

## Self-reported evaluations

### Descriptive stats

```{r}

# overall self-reported evaluations
dat_mean_ratings <- data_processed_after_exclusions |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n()) |>
  mutate(group = "Full sample")

# self-reported evaluations by gender category
dat_mean_ratings_by_gender <- data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n())

# combine both into one table
bind_rows(dat_mean_ratings,
          dat_mean_ratings_by_gender) |>
  select(Subset = group, Mean, SD, n) |> # select variables of interest, and rename one 
  mutate(Subset = stringr::str_to_sentence(Subset)) |> # Change the case of the Subset variable so that it prints nicely
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  add_header_above(header = c(" " = 1, "Self-reported evaluations" = 3)) |>
  kable_classic(full_width = FALSE)

```

### Descriptive plot

```{r}

ggplot(data_processed_after_exclusions, aes(x = mean_evaluation)) +
  geom_histogram(binwidth = 1,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("Mean self-reported evaluation") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 7)) +
  coord_cartesian(xlim = c(1, 7)) +
  theme(panel.grid.minor = element_blank())

```

## AMP evaluations

### Descriptive stats

\TODO add table of means, SDs, Ns

```{r}
data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(AMP_score, na.rm = TRUE),
            SD = sd(AMP_score, na.rm = TRUE),
            n = n()) |> 
  mutate_if(is.numeric, round_half_up, digits = 2) |> 
  kable() |>
  add_header_above(header = c(" " = 1, "AMP Score" = 3)) |>
  kable_classic(full_width = FALSE)


```

### Descriptive plots

```{r}

ggplot(data_processed_after_exclusions, aes(x = AMP_score)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("AMP score") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 10))

```

# Analyses & hypothesis tests

## Cronbachs alpha of like, prefer, positive

```{r}
library(psych)
alpha(subset(data_processed_after_exclusions, 
             select = c(prefer, positive, like)), check.keys =TRUE)
```
Cronbach's alpha for the three scales of self-reported evaluations is .88. This is considered a very good value.

## Self-reported evaluations are correlated with evaluations on the AMP

### Plot

```{r}

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(y = AMP_score,
           x = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  ylab("AMP score") +
  xlab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

```

More complex plots:

#### Axial histograms

Scatter plots with axial histograms using ggExtra: https://cran.r-project.org/web/packages/ggExtra/vignettes/ggExtra.html

\TODO add axial histograms to a scatter plot. Split both the scatter plot and the histograms by gender.


```{r}

p <- ggplot(data_processed_after_exclusions, aes(x = AMP_score,
                                                 y = mean_evaluation, color = gender)) +
  geom_point(alpha = 0.7) + # turning down alpha a bit so you can see where are more points and where are less
  geom_smooth(se = FALSE, method = "lm") + 
  theme_classic() +
  theme(legend.position = c(.15, .7))+
  xlab("AMP Score")+
  ylab("Mean evaluation")
ggMarginal(p, groupColour = TRUE, groupFill = TRUE)


```


Here is a different way, by making plots for each gender and then putting them together.

```{r}


data_processed_after_exclusions_women <- data_processed_after_exclusions |> 
  filter(gender == "female")

p_women <- ggplot(data_processed_after_exclusions_women, 
                  aes(x = AMP_score,
                      y = mean_evaluation)) +
  geom_jitter()+
  ggtitle("female")

p_women <- ggMarginal(p_women, type = "histogram")

data_processed_after_exclusions_men <- data_processed_after_exclusions |> 
  filter(gender == "male")

p_men <- ggplot(data_processed_after_exclusions_men, 
                aes(x = AMP_score,
                    y = mean_evaluation)) +
  geom_jitter() +
  ggtitle("male")

p_men <- ggMarginal(p_men, type = "histogram")

data_processed_after_exclusions_nonbinary <- data_processed_after_exclusions |> 
  filter(gender == "nonbinary")

p_nonbinary <- ggplot(data_processed_after_exclusions_nonbinary, 
                      aes(x = AMP_score,
                          y = mean_evaluation)) +
  geom_jitter() +
  ggtitle("nonbinary")

p_nonbinary <- ggMarginal(p_nonbinary, type = "histogram")

list(p_women, p_men, p_nonbinary) |> 
  wrap_plots()


```

#### Labelled points

Label points using ggrepel: https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html

\TODO Label the points in a scatter plot using their participant codes.

```{r}
p_label_points <- ggplot(data_processed_after_exclusions, 
                         aes(x = AMP_score,
                             y = mean_evaluation, 
                             label = subject)) + 
  geom_jitter() + 
  geom_text_repel(max.overlaps = 100) +
  theme_classic() +
  xlab("AMP Score")+
  ylab("Mean evaluation")

p_label_points

```

#### Magnify areas

Magnify areas of your plot with ggmagnify: https://hughjonesd.github.io/ggmagnify/

\TODO Magnify an area of one of your scatter plots, eg where there are a lot of data points in a small area.


```{r}
p <- ggplot(data_processed_after_exclusions, aes(x = AMP_score,
                                                 y = mean_evaluation, color = gender)) +
  geom_jitter() +
  geom_smooth(se = FALSE, method = "lm") + 
  theme_classic() +
  xlab("AMP Score")+
  ylab("Mean evaluation")

from <- c(xmin = 0.45, xmax = 0.65, ymin = 0.75, ymax = 1.25)
to <- c(xmin = 0, xmax = 0.40, ymin = 4, ymax = 5)

p + geom_magnify(from = from, to = to)

```

### Test

\TODO  run an appropriate test. Below the output, interpret the results: write a few sentences that report and interpret the results following APA reporting style.

```{r}
cor.test(data_processed_after_exclusions$AMP_score, data_processed_after_exclusions$mean_evaluation, method="spearman")

```
Spearman’s rank correlation was computed to assess the relationship between the self-reported evaluations and evaluations on the AMP.
There was no significant correlation between the two variables, r(88) = -0.13, *p* = 0.231.


## Self-reported evaluations differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.

Histogram:

```{r}
# get only female and male in the dataset
data_processed_after_exclusions_without_nonbinary <- subset(data_processed_after_exclusions, gender %in% c("female", "male"))


p_hist <- ggplot(data_processed_after_exclusions_without_nonbinary, aes(x = mean_evaluation, fill = gender)) +
  geom_histogram(binwidth = 1,
                 boundary = 0) +
  labs(title = "Self-reported Evaluations by Gender",
       x = "Mean evaluation",
       y = "Frequency") +
  theme_classic() +
  theme(legend.position = "none")+
  scale_x_continuous(breaks = pretty_breaks(n = 7)) +
  coord_cartesian(xlim = c(1, 7)) +
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~gender, scales = "free")

p_hist

```

Violin Plot:
```{r}
p_violin <- ggplot(data_processed_after_exclusions_without_nonbinary, aes(x = gender, y = mean_evaluation, fill = gender)) +
  geom_violin() +
  labs(title = "Self-reported Evaluations by Gender",
       x = "Gender",
       y = "Mean evaluation") +
  theme_minimal()

p_violin

```
Raincloud Plot:

```{r}

p_raincloud <- ggplot(data_processed_after_exclusions_without_nonbinary, aes(x = gender, y = mean_evaluation, fill = gender)) +
  geom_rain(
    point.args = list(alpha = 0.5)
  ) +
  labs(title = "Self-reported Evaluations by Gender",
       x = "Gender",
       y = "Mean evaluation") +
  theme_classic()+
  scale_y_continuous(
    breaks = seq(1, 7, 1),
    limits = c(1,7)
  )
p_raincloud

```


### Test

\TODO run an appropriate test. Below the output, interpret the results: write a few sentences that report and interpret the results following APA reporting style.

```{r}


t.test(mean_evaluation~gender, data = data_processed_after_exclusions_without_nonbinary, alternative = "two.sided")

```
Welch two sample t-test was computed to assess if there is a difference in self-reported evaluations between men and women.
Men (*M* = 1.81, *SD* = 1.03) had significantly higher values in the self-reported evaluations compared to women (*M* = 1.32, *SD* = 0.71), *t*(83.14) = -2.62, *p* = .01.



## Evaluations on the Affect Misattribution Procedure differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.

This time, vary the labeling and order of the legend, e.g., capitalise "Men" and "Women", and know how to change the order of the factors.

Histogram:

```{r}
data_processed_after_exclusions_without_nonbinary_reordered <- data_processed_after_exclusions_without_nonbinary |> 
  mutate(gender = factor(gender, levels = c("male", "female")))

p_hist2 <- ggplot(data_processed_after_exclusions_without_nonbinary_reordered, aes(x = AMP_score, fill = gender)) +
  geom_histogram(binwidth = 0.1,
                 boundary = 0) +
  scale_fill_discrete(labels=c('Men', 'Women'))+
  labs(title = "AMP Score by Gender",
       x = "Mean evaluation",
       y = "Frequency") +
  theme_classic() +
  theme(legend.position = "none")+
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~gender, scales = "free",
             labeller = labeller(gender = c("male" = "Men", "female" = "Women")))

p_hist2

```
Violin Plot:
```{r}
p_violin2 <- ggplot(data_processed_after_exclusions_without_nonbinary_reordered, aes(x = gender, y = AMP_score, fill = gender)) +
  geom_violin() +
  labs(title = "AMP Score by Gender",
       fill = "Gender",
       x = "Gender",
       y = "AMP Score") +
  scale_fill_discrete(labels=c('Men', 'Women'))+
  scale_x_discrete(labels=c('Men', 'Women'))+
  theme_minimal()

p_violin2

```
Raincloud Plot:

```{r}

p_raincloud2 <- ggplot(data_processed_after_exclusions_without_nonbinary_reordered, aes(x = gender, y = AMP_score, fill = gender)) +
  geom_rain() +
  labs(title = "AMP Score by Gender",
       x = "Gender",
       fill = "Gender",
       y = "AMP Score") +
  scale_fill_discrete(labels=c('Men', 'Women'))+
  scale_x_discrete(labels=c('Men', 'Women'))+
  theme_classic()
p_raincloud2
```


### Test

\TODO run an appropriate test. Below the output, print an interpretation of the results generated by the 'easystats' package [report](https://easystats.github.io/report/). I.e., use `report::report()`.

```{r}


t.test(AMP_score~gender, data = data_processed_after_exclusions_without_nonbinary, var.equal = TRUE, alternative = "two.sided")

```

```{r}
report(t.test(data_processed_after_exclusions_without_nonbinary$AMP_score~data_processed_after_exclusions_without_nonbinary$gender, var.equal = TRUE, alternative = "two.sided"))

```


## Combining plots

Combine plots using patchwork: https://patchwork.data-imaginist.com/

\TODO Combine at least three of the above plots into one.

```{r}

combined_plot <- p_hist2 + p_violin2 + p_raincloud2
combined_plot

```

## Saving plots

Save plots to disk with `ggsave()`

\TODO Save the above combined plot to disk as both .png and .pdf. Ensure the png has at least 300dpi resolution.

```{r}
dir.create("../data/plots/")

ggsave("../data/plots/combined_plot.png", width = 20, plot = combined_plot, dpi = 300, device = "png")

ggsave("../data/plots/combined_plot.pdf", width = 20, plot = combined_plot, dpi = 300, device = "pdf")



```

# Session info

```{r}

sessionInfo()

```



